|

### 2023-08-21 chapter_4/src/prr/dcr_color_and_non-political_behaviors_prr.qmd
a little bit of drafting of the experimental design

To most of us, we likely do not trust this as a basis to infer partisanship as other forms of information such as whether a potential neighbor is displaying a political yard sign. However, these usually only show up every couple of years during the election season and are displayed by those who tend to be highly politically active -- so, they are not universally displayed [@makse_politics_2019]. Additionally, other decisions such as the color of the house are determined by Homeowners Associations (HOAs). The choices of car colors are less constrained by other external factors. So, we may reasonably assume that if we see a neighborhood with a significant number of red cars, we may assume that the local sports team color must be red. However, if we are searching for cues toward the partisan composition of the neighborhood, we may then assume that the neighborhood has a large number of Republicans, whether that is accurate or not.

@@

# Experiments

- Make real estate flyers that are like a conjoint providing information about the neighborhood. Do something with a "favorite car in the neighborhood" and vary whether it is a red or a blue car.

>>>>

To most of us, we likely do not trust the color of a car as a basis to infer partisanship as other forms of information such as whether a potential neighbor is displaying a political yard sign. However, political yard signs usually only show up every couple of years during the election season and are displayed by those who tend to be highly politically active -- so, they are not universally displayed [@makse_politics_2019]. Additionally, other decisions such as the color of the house are determined by Homeowners Associations (HOAs). The choices of car colors are less constrained by other external factors. So, we may reasonably assume that if we see a neighborhood with a significant number of red cars, we may assume that the local sports team color must be red. However, if we are searching for cues toward the partisan composition of the neighborhood, we may then assume that the neighborhood has a large number of Republicans, whether that is accurate or not. Once we have used that association between the colors red and blue with partisan affiliation, we will form a snap-judgment of our neighbors based on the color of their cars when absent other information. That impression will then elicit a positive or negative attitude toward the neighbors and neighborhood as well as encourage a congruent behavioral reaction such as wanting to avoid the neighborhood or an affinity toward it. This impression and behavioral reaction forms pre-consciously and is quickly adjusted with new information from other sources of information. However, the impression constrains what memories are accessible and therefore constrains the attitude expression and behavioral motivations toward objects even when we collect more information that either fits with or goes against our initial impression.

There are many characteristics of a neighborhood besides car color that can better convey information about the partisanship of a neighborhood such as whether it is mixed use, in an urban or rural setting, whether there are Black Lives Matter flags, and whether the cars tend to be pickup trucks or sedans. However, the role of color is to act as a form of pre-conscious impression formation. While all of these alternative, and better, cues towards partisanship may be present, they are more complicated forms of visual information that will require more laborious processing before attitude formation. The point of the snap-judgment model is to illustrate how extremely simple visual information, such as color, shapes our thinking due to the cognitive processes that underlie impression formation and the effects of impressions after particular neurological paths are made "hot".

@@

# Experimental test

Applying the snap-judgment model to an individual's choice to move to a neighborhood yields the following hypotheses: with all else equal, Republicans that see that the modal color of car is Red are more likely to report wanting to move to the neighborhood than if the modal car color were blue ($H_1$); with all else equal, Democrats that see the modal color of Blue are more likely to report wanting to move to the neighborhood than if the modal car color were red ($H_2$); once we include more information about a neighborhood's partisan composition, then the strength of their desire to live in a neighborhood will increase if congruent with their partisan affiliation ($H_3$) and will weaken if incongruent with their partisan affiliation ($H_4$).

To test these hypotheses, I conduct an experiment involving a fictional real estate posting. The treatment involves randomly assigning participants to view one of two versions of a fictional real estate posting. The characteristics of the neighborhood are exactly the same between the two conditions except for the "favorite car" in the neighborhood. In the first condition the "favorite car" is a red SUV. In the second condition, the "favorite car" is the same make and model of SUV as in the first condition, but blue. Participants are then given a scale from 0-10 to rate the neighborhood and are presented with a Likert scale to indicate, "How much would [they] like to live in this neighborhood?" Once the participants have responded to these questions, they are then given a "second page" describing the neighborhood. The information subjects sees depends on which of two conditions they are randomly assigned into. In one condition, the information indicates that the neighborhood is "40 minutes from the nearest metro area" and "60% voted for the Republican candidate in the latest congressional election." The second condition presents the same information but instead describes the neighborhood as "10 minutes from the nearest metro area" and "60% voted for the Democratic candidate in the latest congressional election." With this new information, I ask participants "With this new information, please update your rating of the neighborhood" on the 0-10 scale and are again presented with a Likert scale to indicate, "How much would [they] like to live in this neighborhood?"
